\chapter{Introduction}
\label{chap:intro}
This chapter introduces the Domain Chat Protocol and defines document conventions.


\section{Purpose}
\sectionrule
This document is intended to describe the overall design and on-wire protocol
for Domain Chat Protocol, or DCP.  DCP is designed to be a modern, open protocol
for both group chats and one-to-one messaging.

Unlike IRC\cite{rfc1459}, DCP is designed to centralise management for disparate
chat networks and unify communities together, making communication easier.  DCP
is designed to streamline community administration, and allow multiple networks
to relay the same community messages without needing additional infrastructure.

Shortcomings in multiple instant messaging (IM) protocols, including both
proprietary protocols such as OSCAR and Skype, and openly developed protocols
such as XMPP\cite{xmpp}, plague modern instant messaging.  DCP is designed to
overcome these flaws, enabling simple, rich messaging from any connected device,
including desktop and notebook computers; tablets; phones; game consoles; and
any other device that supports TCP.


\section{Document conventions}
\sectionrule
This document uses formatted text in the following manners:

\begin{itemize}
\item
\texttt{Monospaced text} denotes a literal protocol command or typed input.

\item
\textit{Italicised text} denotes placeholder text that shall be changed to an 
appropriate value.

\item
\textbf{Bold-face text} denotes the definition of a new term.
\end{itemize}

The key words "MUST", "MUST NOT", "SHOULD", "SHOULD NOT", "MAY", and "OPTIONAL"
in this document are to be interpreted as described in RFC 2119\cite{rfc2119}.